---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:default
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")
head(activity)
```

```{r}
str(activity)
```

```{r}
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```


## What is mean total number of steps taken per day?

```{r}
total_step <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
head(total_step)
```

For histogram
```{r}
par(mfrow = c(1, 1))
hist(total_step$steps, breaks = 20, 
     main = "Total Number of Steps Taken Each Day",
     col = "grey", border = "white", xlab = "Step", axes = FALSE)
axis(1)
axis(2, las = 1)
```

Mean and Median of total steps taken
```{r}
mean(total_step$steps)
median(total_step$steps)
```

## What is the average daily activity pattern?

```{r}
avg_step <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
plot(avg_step$interval, avg_step$steps, type = "l", lwd = 2, col = "navy",
     main = "Time Series: Average Number of Steps Taken", axes = FALSE,
     xlab = "5-minute interval", ylab = "Average number of steps")
axis(1)
axis(2, las = 1)
```

Average maximum no. of steps
```{r}
avg_step$interval[which.max(avg_step$steps)]
```

The 835-th 5-minute interval contains the maximum number of steps.

## Imputing missing values

```{r}
sum(is.na(activity))
```

There are 2304 missing values in the dataset. 

2) dataset equal to original dataset but with the missing data filled in .

```{r}
imp <- activity
for (i in avg_step$interval) {
    imp[imp$interval == i & is.na(imp$steps), ]$steps <- 
        avg_step$steps[avg_step$interval == i]
}
head(imp)

sum(is.na(imp))

```

3) Histogram of total number of steps taken each day with mean and median with imputing missing values

```{r}
total_step_imp <- aggregate(steps ~ date, data = imp, sum, na.rm = TRUE)
hist(total_step_imp$steps, breaks = 20, 
     main = "Total Number of Steps Taken Each Day (Imputed)",
     col = "grey", border = "white", xlab = "Step", axes = FALSE)
axis(1)
axis(2, las = 1)
```

```{r}
mean(total_step_imp$steps)
```

```{r}
median(total_step_imp$steps)
```

Mean is the same as the first part of the assignment.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
imp$day <- weekdays(imp$date)
imp$week <- ""
imp[imp$day == "Saturday" | imp$day == "Sunday", ]$week <- "weekend"
imp[!(imp$day == "Saturday" | imp$day == "Sunday"), ]$week <- "weekday"
imp$week <- factor(imp$week)
```
Making a time series plot
```{r}
avg_step_imp <- aggregate(steps ~ interval + week, data = imp, mean)
library(lattice)
xyplot(steps ~ interval | week, data = avg_step_imp, type = "l", lwd = 2,
       layout = c(1, 2), 
       xlab = "5-minute interval", 
       ylab = "Average number of steps",
       main = "Average Number of Steps Taken (across all weekday days or weekend days)")
```





